### 2016-03-08 doc/publications/plos16/plos16.tex
Intro

Important because:

Computational tools such as Apache Spark provide a framework for designing and implementing complex analyses.

In addition of Spark ML, provides popular machine learning libraries. It includes supervised approaches such as Decision trees, and the ensemble Random Forests, as well as unsupervised approaches such as k-means clustering.

We previously demonstrated the versatility and scalability of Spark by developing VariantSpark, a framework to allow users to apply Spark and Spark ML's algorithms to VCF files.

Although we successfully clustered the 2500 individuals from the 1000 Genomes Project phase 3 dataset based variants from the 22 autosomes, this task approached the limits of our cluster.

Although our VCF pre-processing phase succeeds with a minimal resource requirement (3GB per executor) the memory required for k-means clustering skyrockets, requiring 24GB per executor to cluster the entire genome.

The reason for this can be attributed to the "curse of dimensionality". Spark ML is designed to process large numbers of samples rather than a large number of dimensions. As a result, data is partitioned along sample boundaries, with entire samples distributed to worker nodes.
As the sample size grows, Spark can simple schedule more jobs to run to cope with the additional samples.

However with the 1000 Genomes Project data, the dimensionality far exceeds the number of samples. This limits the number of jobs that Spark can create, and instead forces it to create fewer, larger jobs.

Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq.~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

@@

>>>>

In the past few years, whole genome sequence (WGS) data has rapidly been becoming an essential part in the clinical space, where traditional methods are failing to produce results.

WGS has proven itself in discovering rare variants [http://www.sciencedirect.com/science/article/pii/S2352396414000498] and diagnosing unknown genetic disorders [MATT MIGHT],

One challenge with analysing WGS data over, for example, traditional genome-wide association (GWA) data, is developing software and algorithms to deal with the huge amounts of data.
Although the sample size of WGS data-sets may be comparable to GWA, the dimensionality of the data may be many times larger. For example, the 1000 Genome data consists of
approximately 2500 individuals with up to 80 million variants. The uncompressed file sizes are close to a terabyte. Of course, it's not just genomic data that is rapidly growing. Many other
types of data

To manage Developed by Apache, Hadoop is one platform for dealing with large data sets is 
Platforms Fortunately computational tools such as Apache Spark are available as a framework for iteratively dealing with big data.

Spark also includes the machine learning (ML) library, Spark ML, providing popular machine learning libraries. Spark ML includes supervised approaches such as decision trees and the ensemble method, random forests, as well as
unsupervised approaches such as k-means clustering.
We previously demonstrated the versatility and scalability of Spark by developing VariantSpark, a framework allowing users to easily analyse Variant Call Format (VCF) files using ML algorithms on the Spark framework. Using VariantSpark,
we successfully built a k-means model on the 2500 individuals from the 1000 Genomes Project, based on their genomic data. We also investigated classifying the samples using Random Forests. We chose random forests due to its
ability to capture complex interactions between features. Also, due to it being an ensemble method where trees can be formed independently, which, is ideal for parallelisation. Furthermore, the ensemble of weak learners can help to minimise overfitting and variance.

However, we observe that Spark ML's random forests implementation isn't able to handle the 1000 Genomes data, and is unable to build a model. The reason for the failure can be attributed to the "curse of dimensionality". Where Spark ML
is designed to process a huge number of samples with a modest dimensionality, the 1000 Genomes data instead contains a huge number of dimensions (features). As a result, the default random forests implementation is unable to partition the data into small-enough tasks
for nodes on the computer cluster to handle.

Therefore, we develop our own 'wide' version of these ML algorithms.
Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq.~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

@@

\subsection*{}
To

\subsection*{CursedForest copes with the curse of dimensionality}
To

\subsection*{CursedForest copes with the curse of dimensionality}
To
